A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies

### 2015-08-13 Manuscript/Revision/response.tex

\documentclass[12pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[osf]{mathpazo}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\signature{Natalie Cooper, Gavin Thomas \\ Chris Venditti, Andrew Meade, \\ and Rob Freckleton}
\address{}
\longindentation=0pt
\begin{document}

@@

We thank the reviewer for their positive comments.

\textcolor{blue}{However, while I understand the authors’ frustrations with the fitting and interpretations of single peak OU processes that are common in the literature, I think they miss an opportunity to really emphasize their concerns here and demonstrate to readers that it's easy to correct oneself when support for OU comes back in comparative analyses. Though the authors do note that a best fitting OU model could be nothing more than a noisy BM processes based on parameter estimates, a closer look at their simulation results reveals that this is also the case here. For example, on their simulated trees rescaled to unit height, the highest median alpha estimated for the BM datasets is 0.165 for the 25 tip Yule tree. This translates into a phylogenetic half life of 4.2 time units however, 4 times the age of the simulated clade, which no sane macroevolutionary biologist would interpret as being too different from Brownian motion at this time scale! Furthermore, the upper 95\% HPDs (Quantiles surely?) are, in many cases low enough to result in questionable acceptance of OU over BM for any of the simulations.}

\textcolor{blue}{I think the authors could do a better job of emphasizing the implications of OU model parameters on the expected outcomes of the process with respect to their own simulations, perhaps by adding a section after line 211 on page 9 that explores alpha estimates and their associated phylogenetic half-lives. If the aim here is to caution readers that a best-fitting OU process does not always translate to evolution about a constrained peak, then showing this with parameter estimates from the simulations would be a useful contribution. This kind of constructive comment would also help those readers who would view this only as a critique of OU models.}

We agree that a half-life approximately 4 times the height of the tree is no different from Brownian motion in any meaningful biological sense. This is exactly the point. It is not uncommon to see interpretations of models based only on model fit and not on the the model parameters hence our emphasis is on problems with model fitting. However, the point about parameter interpretation is equally important and we have added a new paragraph (lines 273-282).
to discuss this explicitly.

@@

\textcolor{blue}{lines 404-407: This works the other way too; not incorporating measurement error could lead one to reject a “high phylogenetic signal” model like EB in favor of BM when EB is intact the true model.}

We have added a comment in the text to say that measurement error can also favour BM over the true model.

\textcolor{blue}{Figure 2 legend what are the black and red dashed lines?}

We have added details to this Figure caption. See our response to reviewer 1 above.

We hope we have responded to all comments appropriately,
